Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-03-04 paper/ms.tex
revised discussion, added refs to Iben & Bodenheimer

\subsection{Magnetic Sequence Turn-off}
\label{sec:dynamo}
% F --> G type members are bright (V < 10 mag) and could be accessible to ZDI. Perhaps we'd see changes in magnetic strengths or topologies stretching from late-G to late-F.

{\bf Figure~\ref{fig:ages} shows that a 10 Myr magnetic model isochrones traces a 5 Myr standard model isochrone for effective temperatures below about 5\,000~K, but traces a 10 Myr standard model isochrone above 6\,000~K. This suggests that magnetic inhibition plays an important role in dictating the structure of cool low-mass stars, but the structure of warmer high-mass stars are relatively insensitive to the influence of magnetic fields. In between these two regimes, the models predict a that a transition occurs in which stars appear to exhibit an increasing sensitivity to magnetic inhibition of convection as effective temperature decreases. }

{\bf As was noted in Sect.~\ref{sec:results}, Figs.~\ref{fig:empirical} and \ref{fig:ages}a indicate that a similar transition occurs over roughly the same effective temperature domain in the empirical HR diagram of Upper Scorpius. Early- to mid-G stars in Upper Scorpius (\teff$ \sim 6\,000$~K) are best described by a 10 Myr standard stellar model isochrone. Instead of continuing to follow a 10 Myr standard model isochrone, late-G stars appear to be better represented by an isochrone intermediate between 5 and 10 Myr causing the empirical sequence to exhibit a shallower slope than standard model predictions between $5\,750 > T_{\rm eff} / {\rm K} > 4\,500$~K. 
That the shallower slope in the empirical HR diagram and the transition predicted by magnetic stellar models are coincident, suggests that the two phenomena may be connected.
}

@@

{\bf The transition exhibited by theoretical magnetic models can be explained by the rapid thinning of outer convective envelopes in stars with masses above about 1.2 $M_{\odot}$.
Once a star reaches an effective temperature of 6\,000 K, it has either a very small or non-existent convective envelope. This provides a natural explanation the convergence of the 10 Myr magnetic isochrone with the 10 Myr standard model isochrone that occurs near this effective temperature in Fig.~\ref{fig:ages}(b).

Assuming that a change in slope of the empirical HR diagram is observed at 4\,500 K, 
Comparison of the empirical HR diagram to magnetic stellar models suggests that the transition's effective temperature domain begins somewhere between 4\,500~K and 5\,000~K, consistent with the knee observed in the empirical data at \teff$\sim 4\,500$~K. According to magnetic stellar models, these effective temperatures correspond to stars with masses between 1.26 \msun\ and 1.40 \msun. This is significant because stars in this mass regime are expected to have relatively thin convective envelopes that eventually disappear entirely. 
}

{\bf % Not only for stars with M ~ 1.2... all intermediate mass stars above 0.28 Msun. 
Convective envelopes in stars with masses greater than $\sim1.2$ $M_{\odot}$ gradually recede from the deep interior toward the stellar surface in response to the ignition of the proton-proton ($p$-$p$) chain, continued release of gravitational potential energy as the star contracts, and a decreasing radiative opacity due to rising gas temperatures \citep{Henyey1955}. During this time, the star's radius is predicted to increase. Eventually, the radius expansion is halted and the convective envelope disappears entirely following ignition of the CN-cycle. Prior to ignition of the CN-cycle, a thinning of the convective envelope means magnetic fields of a given strength will have a larger impact on the thermal structure of the super-adiabatic region. However, a restructuring of the super-adiabatic layers in a star with a thin convective envelope has less of an impact on the thermal structure of the deeper underlying layers, which are assumed to be in radiative equilibrium, than for a star whose interior is largely convective. The thinning of the convective envelope is also accompanied by a drop in the surface gas pressure as the star's radius inflates thereby lowering the permitted equipartition magnetic field strength. The combination of these two effects suggests that the transition from a magnetic sequence to a non-magnetic sequence should coincide with the disappearance of the convective envelope.} % confirm discussion of super-adiabatic layers

>>>>

\subsection{Magnetic Field Evolution \& Radiative Core Growth}
\label{sec:dynamo}
% F --> G type members are bright (V < 10 mag) and could be accessible to ZDI. Perhaps we'd see changes in magnetic strengths or topologies stretching from late-G to late-F.

{\bf Figure~\ref{fig:ages} shows that a 10 Myr magnetic model isochrones traces a 5 Myr standard model isochrone for effective temperatures below about 5\,000~K, but traces a 10 Myr standard model isochrone above 6\,000~K. This suggests that magnetic inhibition plays an important role in dictating the structure of cool low-mass stars, but the structure of warmer high-mass stars is relatively insensitive to the influence of magnetic fields. In between these two regimes, the models predict a transition in which stars appear to exhibit an increasing sensitivity to magnetic inhibition of convection as effective temperature decreases. }

{\bf As was noted in Sect.~\ref{sec:results}, Figs.~\ref{fig:empirical} and \ref{fig:ages}a indicate that a similar transition occurs over roughly the same effective temperature domain in the empirical HR diagram of Upper Scorpius. Early- to mid-G stars in Upper Scorpius (\teff$ \sim 6\,000$~K) are best described by a 10 Myr standard stellar model isochrone. Instead of continuing to follow a 10 Myr standard model isochrone, late-G stars appear to be better represented by an isochrone intermediate between 5 and 10 Myr causing the empirical sequence to exhibit a shallower slope than standard model predictions between $5\,750 > T_{\rm eff} / {\rm K} > 4\,500$. 
That the shallower slope in the empirical HR diagram and the transition predicted by magnetic stellar models are coincident, suggests that the two phenomena may be connected.
}

@@

{\bf The transition exhibited by theoretical magnetic models can be explained by a rapid thinning of outer convective envelopes in stars with masses above about 1.2 $M_{\odot}$. Once a star reaches an effective temperature of 6\,000 K, it has a very small or non-existent convective envelope. In the absence of a convective envelope, magnetic inhibition of convection is irrelevant. Therefore, any differences observed between magnetic and standard stellar model isochrones at this effective temperature are the result of the magnetic field's influence on stellar structure at earlier evolutionary phases. This provides a natural explanation for the convergence of the 10 Myr magnetic isochrone with the 10 Myr standard model isochrone that occurs near this effective temperature in Fig.~\ref{fig:ages}(b). 
It then seems reasonable to posit that the shallow slope of the empirical HR diagram and the observed transition in the magnetic models trace the demise of outer convective envelopes and the growth of radiative cores in young, pre-main-sequence stars.}

{\bf % Not only for stars with M ~ 1.2... all intermediate mass stars above 0.28 Msun. 
Stars with masses above about $0.28\ M_{\odot}$ are predicted to, at least temporarily, form a small radiative core during their pre-main-sequence contraction. Release of gravitational potential energy as stars contract heats the gas causing radiative opacities to decrease and igniting fusion from the proton-proton ($p$-$p$) chain \citep{Henyey1955}. Stars with masses below about $1.2\ M_{\odot}$ continue to contract until energy production from the $p$-$p$ chain is sufficient to stabilize the core (and thus the star) against further contraction. However, above $1.2\ M_{\odot}$, core temperatures become sufficiently hot enough to produce hydrogen burning via the CN-cycle at an appreciable level. Increased energy production from the CN-cycle, while initially insufficient to support the core against gravitational contraction, causes the radiative core to continue growing and the convective envelope to continue shrinking. Increasing energy production from the CN-cycle and $p$-$p$ chain causes a star's outer layers are predicted to expand while the core's contraction slows. Eventually, the CN-cycle reaches equilibrium, a convective core develops, the radius expansion is reversed, and the convective envelope disappears entirely \citep{Iben1965,Bodenheimer1965}. }

{\bf 
To fully understand the model-predicted transition in Fig.~\ref{fig:ages}(a) and the reasons why magnetic models do not entirely reproduce the morphology of the empirical transition, it is important to consider how the magnetic field evolves as a star's convective envelope shrinks.
Prior to ignition of the CN-cycle, a thinning of the convective envelope means magnetic fields of a given strength will have a larger impact on the thermal structure of the super-adiabatic region. However, a restructuring of the super-adiabatic layers in a star with a thin convective envelope has less of an impact on the thermal structure of the deeper underlying layers, which are assumed to be in radiative equilibrium, than for a star whose interior is largely convective. The thinning of the convective envelope is also accompanied by a drop in the surface gas pressure as the star's radius inflates thereby lowering the permitted equipartition magnetic field strength. The combination of these two effects suggests that the transition from a magnetic sequence to a non-magnetic sequence should coincide with the disappearance of the convective envelope.} % confirm discussion of super-adiabatic layers

{\bf Assuming that a change in slope of the empirical HR diagram is observed at 4\,500 K, 
Comparison of the empirical HR diagram to magnetic stellar models suggests that the transition's effective temperature domain begins somewhere between 4\,500~K and 5\,000~K, consistent with the knee observed in the empirical data at \teff$\sim 4\,500$~K. According to magnetic stellar models, these effective temperatures correspond to stars with masses between 1.26 \msun\ and 1.40 \msun. This is significant because stars in this mass regime are expected to have relatively thin convective envelopes that eventually disappear entirely. 
}
